---
title: "03 -- Selection Syntax"
author: Sebastian Raschka
output:
  github_document:
    fig_width: 5
    toc: true
    toc_depth: 2
---

Source file: https://github.com/rasbt/R-notes/blob/master/03-selection-syntax.Rmd

# Selection Syntax


## Selecting Elements from Vectors and Matrices (Single Bracket Indexing)

- We can use the single brackets `[]` indexing approach to return one or more elements from a vector or data frame, not that it returns the same object as the original.

```{r}
vec_a <- c("x", "y", "z")

vec_a[1] # return a vector containing the first element

vec_a[1:2] # return a vector containing the first two elements
```

- We can also combine the single bracket selection approach with logical operators.
- Suppose we have the following integer vector; we can use the `>=` operator to select all numbers that are equal to or larger than a certain value (here: greater than or equal to 5):

```{r}
vec_b <- c(1:9)
vec_b

vec_b[vec_b >= 5] 
```

- Also, we can chain multiple logical operators; let's select all elements greater than 5 and smaller than 8:

```{r}
vec_b[(vec_b > 5) & (vec_b < 8)]
```


### How about Matrices?

- Analogous to the vector examples above, we can also use the single bracket notation to select elements and entire rows and columns from matrices. Let's walk through some examples using a simple 3x2 matrix:

```{r}
my_matrix <- matrix(1:6, 3, 2) # 3x2 matrix
my_matrix
```

- Use two values separated by a comma; the first value is the row index, the second value is the column index.
- E.g., to select the 1st element in the 2nd row, we can do

```{r}
my_matrix[2, 1]
```

- To select an entire row, we omit the value after the comma. E.g., to select the 2nd row:

```{r}
my_matrix[2, ]
```

- Similarly, we can select an entire column by leaving the row value empty. E.g., to select the 2nd column:

```{r}
my_matrix[, 2]
```

- We can also select with logical vectors and matrices. Suppose we have the following matrix with missing values:

```{r}
my_matrix_with_na = my_matrix # make a copy

my_matrix_with_na[1, 2] <- NA # overwrite element with NA value
my_matrix_with_na
```

- We can create a logical matrix, a mask, denoting wether a value is a `NA` or not:

```{r}
mask <- is.na(my_matrix_with_na)
mask
```

- The reason why we call this mask is that it allows us to select us specific values from the original matrix, i.e., all the TRUE values:

```{r}
my_matrix_with_na[mask]
```

- Note that we can also invert the selection using the `!` operator (which is more useful in this case, assuming we want to filter out NA values). E.g., we can select all the non-NA values as follows:

```{r}
!mask # inverted mask 
```

```{r}
my_matrix_with_na[!mask]
```

## Selecting from Data Frames

- Suppose we have the following data frame:

```{r}
df <- data.frame(MyIntegerVar = 1:4, MyCharVar = c("A", "B", "A", "B"), MyBoolVar = c(T, F, T, T))
df
```

- To get the 3rd row, we can use the same bracket notation as above for matrices:

```{r}
df[3, ]
```

- Similarly, we can select the 3rd column as follows:

```{r}
df[, 3] # returns a vector

# or

df[3] # returns a data frame with 1 column
```

### Different Ways for Selecting Columns from a Data Frame

- However, when working with data frames, we can take advantage of the column names when selecting columns. E.g., to select the 3rd column similar to `df[3]`, we can do the following:

```{r}
df['MyBoolVar']
```

- And to select the third column similar to `df[, 3]` -- returning a vector instead of a 1-column data frame -- we can use the `$` symbol as follows:

```{r}
df$MyBoolVar
```


### Selecting Elements from a Data Frame

- Note that the selection syntax works similar to a matrix; to select the 1st element of the 2nd row, we can do

```{r}
df[2, 1]
```
- however, we can also use the column name approach:

```{r}
df[2, 'MyIntegerVar']
```
or

```{r}
df$MyIntegerVar[2]
```

### Conditional Selection

- Also, similar to vectors and matrices, we can use conditional statements to select values. For instance, we can use the following to select all entries for which `MyIntegerVar` is greater than 1 and `MyBoolVar` is true

```{r}
df[(df['MyIntegerVar'] > 1) & (df['MyBoolVar'] == TRUE), ] 
```