---
title: "07 -- Plotting"
author: "Sebastian Raschka"
date: "6/17/2020"
output:
  github_document:
    fig_width: 5
    toc: true
    toc_depth: 2
---

Source file: https://github.com/rasbt/R-notes/blob/master/07-plotting.Rmd




# Plotting

- There are several different plotting libraries available in R. Here, we are going to focus on the built-in plotting capabilities (interested users may want to take a look at `ggplot2`, which is a popular plotting library some users prefer. You can find more information about it at https://ggplot2.tidyverse.org)

## Stem-and-Leaf plots

- The example dataset used here contains information about the bark and concealment distances (in meters) of rural and urban prairie dogs:

- Stem-and-leaf plots ar probably the simplest type of plot, and we could construct it conveniently by hand for small dataset if we had to.
- using the `stem` function, we can construct the stem-and-leaf plot in R as followes.


```{r}
urban_barkdist <- c(29, 10, 15, 41, 18, 18, 12, 45, 34, 30, 22, 26, 18)
urban_concealdist <- c(8, 8, 10, 8, 9, 10, 9, 10, 9, 6, 6, 7, 5)
rural_barkdist <- c(40, 47, 38, 59, 45, 52, 57, 50, 50, 49, 50, 43)
rural_concealdist <- c(14, 12, 24, 30, 37, 46, 39, 22, 38, 38, 20, 33)
```

```{r}
stem(urban_barkdist)
```

- The left most digit (before the "|") is the first digit of the values in the dataset. The values next to the stem represent the second digit of the values shown there. in the case of the `urban_barkdist` variable, we have 10, 12, 15, 18, 18, 18 (sorted in ascending order) for the first row, which become `1 | 025888`
- This type of plot provides us with a quick view of the distribution of the data. In the example above, we have a right-skewed distribution (the tail is to the right.)

## Histograms

```{r}
hist(urban_barkdist)
```
```{r}
hist(urban_barkdist, breaks=seq(0, 50, by = 5))
```

## Barplot

```{r}
barplot(c(mean(urban_barkdist), mean(rural_barkdist)),
        xlab="Bark distance (m)",
        names.arg=c("Urban", "Rural"),
        ylim=c(0, 65))
```

## Scatterplots

```{r}
plot(urban_barkdist, urban_concealdist,
     xlab="Bark distance (m)",
     ylab="Concealment discance (m)")
```

```{r}
plot(urban_barkdist, urban_concealdist,
     xlim=range(1, 60), 
     ylim=range(1, 60),
     xlab='Bark distance (m)', 
     ylab='Concealment distance (m)');
points(rural_barkdist, rural_concealdist, pch=2)
legend(1, 50, pch=1:2, legend=c("Urban", "Rural"))
```

## Boxplots
